Air cargo load and route planning in pickup and delivery operations

### 2024-02-06 latex/ESWA_carlos.tex
Update

\begin{abstract}
In the aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of load imbalance due to the urgency required for loading, immediate take-off, and mission accomplishment. Transport planners deal with trip itineraries, prioritisation of items, building up pallets, and balanced loading, but there are no commercially available systems that can integrally assist in all these requirements. This enables other risks, such as improper delivery, excessive fuel burn, and possible safety issues due to cargo imbalance, as well as a longer than necessary turn-around time. This {\it NP-hard}\/ problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modelled using standardised pallets in fixed positions. We developed a strategy to solve this problem, considering historical transport data from some Brazilian hub networks, and performed several experiments with a commercial solver, {\color{blue}five} known meta-heuristics, and a new heuristic designed specifically for this problem. By using a portable computer, our strategy quickly found practical solutions to a wide range of real problems in much less than operationally acceptable time.
\end{abstract}

@@

The problem addressed in this work is the definition of the route of an aircraft that loads and unloads hundreds of items in up to seven hubs, dealing with weight, volume, and balance constraints, and maximising the benefit-cost ratio. This challenge has become even more acute during the COVID-19 pandemic due to the need for urgent medical supplies. Delays in shipping essential items such as respirators to critical areas highlighted the need for optimised solutions.

@@

@@

		\caption{Brazilian airports distances ($km$)} \label{tab:costs}
		\centering
		\footnotesize

@@

In all these works, there is a great diversity of scenarios and solutions: some consider items in two dimensions, and others in three dimensions; some used integer programming, and others developed specific heuristics. Among the most recent, we can highlight:
\begin{itemize}

@@

	\item \cite{zhao2023} proposed three integer linear programming models: a bi-objective optimisation model (BOM), a combinatorial optimisation model (COM), and an improved combinatorial optimisation model (IOM). Its objectives were the maximum possible loading and the smallest CG deviation. Four scenarios with multiple metrics are resolved for the {\it Boeing 777F}. The results showed that the BOM has the fastest solution speed, but the CG deviation is the worst; in several cases, the CG deviation is unacceptable. COM has the longest solution time, which is difficult to tolerate in practice.

@@

Throughout this text, we call {\it packed content}\/ a set of items of the same destination stacked on a pallet and covered with a restraining net. It is considered a single item, having the same attributes as its components, whose values are the sum of individual scores, weights, and volumes. To ensure accuracy in pickup and delivery operations, packed content must remain on board until its destination.

@@

\section{Solution strategies}
\label{algorithms}

Throughout our research, we have thoughtfully described ACLP+RPDP in standard MIP format and found that no solver can handle its practical cases in a feasible time. Thus, as ACLP+RPDP is highly complex, involving four intractable and interconnected sub-problems, we decided to focus only on {\it real cases}, developing quick node-by-node solutions, not necessarily optimal, but which would allow us to obtain a complete tour.

@@

			\node [mybox, below=0.7cm of one] (two) { \small {\bf Number of tested tours:}\\
				\small \ \ \ {{\bf if} $K \leq 6$}\\
				\small \ \ \ \ \ \ {$ntours = 2$ \ \ (optimal TSP solutions)}\\
				\small \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\it or}\\				
				\small \ \ \ \ \ \ {$ntours = K!$ \ \ (all possible tours)}\\
				\small \ \ \ {\bf else}\\
				\small \ \ \ \ \ \ {$ntours = 100$ \ \ (best tours obtained with a TSP heuristic)}
			};

@@

\begin{algorithm}[htp]

	\caption{Solving the ACLP+RPDP}  \label{alg:main}

@@

				\For {$x \gets 1$ to $ntours$} \label{main:loop1}

					\State $f(S_K[x]) \gets SolveTour(S_K[x], L, M, C, N, method, tmax/ntours)$ \label{main:fpi}

@@

				\State $answer[scenario,surplus,method] \gets \max f(S_K[x])$ \label{main:a2}
			\EndFor
	\end{algorithmic}

@@

{\color{red}$\pi_k$ are tours} {\color{blue}$\pi$ are tours}, that is, permutations with the $K$\/ nodes. When $ntours=2$, $\pi_1$\/ and $\pi_2$\/ are the optimal solutions of the corresponding TSP. When $ntours=K!$, $\pi_k$\/ is the $k^{th}$ permutation of $S_K$. Finally, when $ntours=100$, $\pi_1, \ldots, \pi_{100}$\/ are the best TSP solutions provided by a heuristic.

@@

\begin{algorithm}[H]
	\caption{Solving the tour $\pi$ with $method$}  \label{alg:tour}
	\begin{algorithmic}[1]

@@

When the aircraft is at the base, the initial graph $G_1$\/ is empty, and there are no packed contents {\color{blue}(line \ref{tour:g11})}. Otherwise, $UpdatePacked$\/ (line \ref{tour:dest}) returns the set of packed contents that have not yet reached their destination and remain on board, rearranging them on the pallets to minimize CG deviation. This allocation is stored in graph $G_1$\/ (line \ref{tour:g12}).

@@

\begin{algorithm}[H]
	\caption{Updating the packed contents that remain boarded at node $\pi_k$}  \label{alg:cons}
	\begin{algorithmic}[1]

@@

\begin{algorithm}[H]
	\caption{Setting pallets destination based on the items to be embarked at node $\pi_k$}  \label{alg:dest}
	\begin{algorithmic}[1]

		\State {\bf SetPalletsDestinations} \,\, {\it in: $M, \pi_k$} \, \, {\it out: $M$}

		\State $vol_x \gets  0$, $x \in L_{\pi_k}$  \label{dest:vector1}

		\State $max \gets 0$  \Comment{destination with maximum volume demand}

		\State $total \gets 0$

		\If {$to_j \in L_{\pi_k}$}

		\State $vol_{to_j} \gets vol_{to_j} + v_j$ 
		\State $total \gets total + v_j$

		\If {$vol_{to_j} > vol_{max}$}

		\State $max \gets to_j$ \label{dest:max} 
		\EndIf

		\EndIf

		\For{$x \in L_{\pi_k}$} \label{dest:propor1}

		\If {$vol_{x} \neq 0$}

		\State $needed \gets \max \{1, \lfloor{ (m-m_{\pi_k}) \times vol_{x}/total}\rfloor \}$  
		\label{dest:needed}
		\State $np \gets 0$

		\For{$i \gets 1$ to $m$}

		\If{($np < needed$) {\bf and} ($T_i^{\pi_k} = -1$)}  
		\State $T_i^{\pi_k} \gets x$
		\State $np \gets np + 1$ \Comment{number of necessary pallets to node $x$}
		\EndIf

		\EndFor
		\EndIf

		\EndFor \label{dest:propor2}

		\For{$i \gets 1$ to $m$} \label{dest:max1}
		\If{$T_i^{\pi_k} \gets -1$} 
		\State $T_i^{\pi_k} \gets max$  \Comment{any remaining pallet is assigned to the maximum demand destination}
		\EndIf
		\EndFor \label{dest:max2}

@@

\begin{algorithm}[H]
	\caption{Cargo score and aircraft torque}  \label{alg:eval}
	\begin{algorithmic}[1]

@@

In this way, the mathematical model for $SolveNode(MIP,\pi_k,G, tmax)$\/ becomes simpler, which finds an allocation of available items at the node $\pi_k$\/ using previously defined values of $L_{\pi_k}$, $T_i^{\pi_k}$, and $a^{\pi_k}_q$. Thus, we use a MIP solver with a runtime limit $tmax$\/ at the node $\pi_k$ to maximize the objective function (\ref{eq:obj2}) with the calculus equations (\ref{eq:newf}) to (\ref{eq:costs2}), subject to the constraints (\ref{eq:tau2}) to (\ref{eq:ifB}). The binary variables $X_{ij}$\/ and $Y_{iq}$\/ define the sets of edges $E_{N_{\pi_k}}$\/ and $E_{Q_{\pi_k}}$, respectively, included in graph $G$.

@@

The selection of edges for $E_{ N_{\pi_k} }$ uses the {\it edge attractiveness}\/ $\theta_{ij}$\/ (\ref{eq:edge}), which can be understood as the tendency to allocate the item $j$\/ to the pallet $i$\/ at the node $\pi_k$. It is directly proportional to the score, and inversely to the volume and the torque of each item.

@@

Considering only the items that can be shipped at node $\pi_k$, Figure \ref{fig:whip} represents $n_{\pi_k}$\/ possible edges $e_{ij}$\/ of the pallet $i$\/ sorted by $\theta_{ij}$\/ in non-ascending order. Initially, {\it Shims}\/ builds a greedy solution for the pallet $i$\/ selecting edges up to index $\eta_1$ ({\it greedy phase}). Then, with the edges between $\eta_1$\/ and $\eta_2$, it elaborates different possible complements ({\it composition phase}), including later the best ones in the same pallet ({\it selection phase}). {\it Shims}\/ is depicted in Algorithm \ref{alg:shims}.

@@

		\State Let $E$\/ be an array of $n_{\pi_k}$ possibles edges of the pallet $i$\/ sorted by $\theta_{ij}$\/ in non-ascending order \label{shims:possible}
		\State $\eta_1 \gets 1$  
		\Repeat
		\State $e_{ij} \gets E_{\eta_1}$

@@

$tmax$\/ is the runtime limit for {\it Shims}, and $level_1$\/ and $level_2$\/ are volume thresholds for the indices $\eta_1$ and $\eta_2$, respectively.

Initially, $Q_{\pi_k}$\/ (line \ref{shims:initQ}) corresponds to the packed contents that remain on board. It is important to remember that $E_{Q_{\pi_k}}$\/ and $M$\/ were modified by the procedure $UpdatePacked(M, Q_{\pi_k}, \pi_k)$ and the procedure $SetPalletsDestinations(M, \pi_k)$. Then, the pallets $i$\/ are considered in non-descending order of $|D_i^{long}|$.

@@

	\item In the {\it greedy phase}\/ (lines \ref{shims:pallets}-\ref{shims:endfirst}), a partial solution for each pallet $i$\/ is constructed by adding edges following $\theta_{ij}$\/ order. The indices $\eta_1$\/ and $\eta_2$\/ refer to the accumulated volumes $V_i \times level_1$\/ and $V_i \times level_2$, respectively, which were defined empirically by the {\it irace}\/ tool \citep{LopezIbanezManuel2016}. We will explain this in detail in the subsection \ref{items}.

@@

{\color{blue}
\subsection{Complexity analysis}

These are the steps to solve the ACLP+RPDP with the complexity analysis:

\subsubsection{Input data}

Load $K$ lines of text to compose a distances matrix for $K$ airports, the $m$ packed contents from previous nodes, and $n$ items parameters from text files are of complexities $O(K)$, $O(m)$, and $O(n)$, respectively:

	\begin{itemize}
		\item $O(K + m + n)$
	\end{itemize}

\subsubsection{Number of tested tours ($T$)}

\begin{itemize}
\item If $K <= 6$, determine the number of tours through permutation. To get permutation of nodes as tours has a complexity of $O(T = K! \cdot K)$, as the number of nodes is permutated (factorial) to generate tours.

	\begin{itemize}
	\item $ O(K + m + n) + O(T = K! \cdot K)$ or
	\item $ O(K + m + n) + O(T = 2)$, if the we are solving the two optimal TSP tours.
	\end{itemize}

\item If $K > 6$, we obtain 100 tours from a heuristic TSP solution (see its time complexity analysis in subsubsection \ref{ga_tsp}):

	\begin{itemize}
		\item $O(K + m + n) + O(T = 100)$
	\end{itemize}
\end{itemize}

\subsubsection{Solving $T$ tours:}

\begin{itemize}
	\item $O(K + m + n) + O(T)$
\end{itemize}

a) each tour with $K$ nodes:
\begin{itemize}
	\item $O(K + m + n) + O(T \cdot K)$
\end{itemize}

b) in each node we have to reset (to eliminate the risk of previous destinations assigned to a pallet) $O(K \cdot m)$ and set the empty pallets destinations: $O(K \cdot m + K \cdot m) = O(2 \cdot K \cdot m) = O(K \cdot m)$
\begin{itemize}
	\item $O(K + m + n) + O(T \cdot K \cdot (K \cdot m))$
\end{itemize}

c) To calculate the time limit to solve each node, it is necessary to determine the potential total volume for each node.: $O(K \cdot n)$
\begin{itemize}
	\item $O(K + m + n) + O(T \cdot K \cdot (K \cdot m + K \cdot n))$
\end{itemize}

d) The time complexity to put the packet contents destined to the next nodes in the tour on board: $O(m)$.
\begin{itemize}
	\item $O(K + m + n) + O(T \cdot K \cdot (K \cdot m + K \cdot n + m))$
\end{itemize}

d) The time complexity to optimize the pallets positions to minimize the CG deviation: $O(m^2)$ ($m$ pallets in $m$ positions)
\begin{itemize}
	\item $O(K + m + n) + O(T \cdot K \cdot (K \cdot m + K \cdot n + m + m^2))$
\end{itemize}

e) Plus the time complexity of the method used to solve an ACLPP with the aircraft partially loaded: $O(C)$
\begin{itemize}
	\item $O(K + m + n) + O(T \cdot K \cdot (K \cdot m + K \cdot n + m + m^2 + C))$
\end{itemize}

\subsubsection{Considering the Shims algorithm: $O(C) = O(Shims)$}

a) the time complexity for sorting the pallets in non-descending order of distances to the CG:
\begin{itemize}
	\item $O(Shims) = m \cdot log(m)$
\end{itemize}

b) As each pallet is solved at a time, the overall Shims complexity is multiplied by $m$.
\begin{itemize}
	\item $O(Shims) = m.(m \cdot log(m))$
\end{itemize}

c) The greedy phase has $O(n)$.
\begin{itemize}
	\item $O(Shims) = m \cdot (m \cdot log(m) + n)$
\end{itemize}

d) The composition phase has $O(n)$ added to the First-Fit Decreasing algorithm time complexity $O(m \cdot log(m))$
\begin{itemize}
	\item $O(Shims) = m \cdot (m \cdot log(m) + n + n + m \cdot log(m))$
	\item $O(Shims) = m \cdot (2n + 2 \cdot m \cdot log(m)) = m \cdot (n + m \cdot log(m))$
\end{itemize}

The overall time complexity to solve the ACLP+RPDP with Shims is:
\begin{itemize}
	\item $O(K+m+n) + O(T \cdot K \cdot (K \cdot m + K \cdot n + m + m^2 + m \cdot (n + m \cdot log(m))))$
\end{itemize}

The overall time complexity provided combines two distinct parts: $O(K+m+n)$ and $O(T \cdot K \cdot (K \cdot m + K \cdot n + m + m^2 + m \cdot (n + m \cdot \log(m))))$. Analysing each part separately and then considering their combined implications provides insight into the algorithm's performance characteristics.

\subsubsection{Complexity analysis considerations}

\textbf{First Part:} $O(K+m+n)$

This component of the time complexity suggests a linear dependency on the sizes of three separate inputs or parameters, denoted by $K$, $m$, and $n$.
These are sequential steps, each influenced by each of these parameters size. This implies that as any of $K$, $m$, or $n$ increases, the time required by the algorithm increases linearly with respect to the largest of these parameters, the number of items to be transported ($n$).

\textbf{Second Part:} $O(T \cdot K \cdot (K \cdot m + K \cdot n + m + m^2 + m \cdot (n + m \cdot \log(m))))$

This portion is considerably more complex, indicating an algorithm with nested dependencies on several variables ($T$, $K$, $m$, and $n$):

\begin{itemize}
	\item $T \cdot K$: repeated iterations $T$ for each of $K$ nodes, nested loops where inner operations are contingent upon $K$.
	\item $K \cdot m + K \cdot n$: These operations scale with both $K$ and each of $m$ and $n$, indicating processes that involve all pairs of $K$ with $m$ and $n$.
	\item $m + m^2$: operations with linear and quadratic dependencies on the number of pallets $m$, indicating that as $m$ grows, the complexity increases more significantly due to the quadratic term $m^2$.
	\item $m \cdot (n + m \cdot \log(m))$: Combines linear scaling with $m$ and $n$ and a logarithmic term $m \cdot \log(m)$, because the algorithm uses divide-and-conquer strategies on $m$ elements, the Shims heuristics.
\end{itemize}

\textbf{Combined Implications}

When combining these parts, the linear terms in the first part ($O(K+m+n)$) are overshadowed by the more complex and potentially rapidly increasing terms of the second part, especially due to the presence of $m^2$ and the product $T \cdot K \cdot (...)$. The quadratic ($m^2$) and logarithmic ($m \cdot \log(m)$) terms indicate that the algorithm's efficiency decreases significantly as the number of pallets $m$ increases.

The overall complexity suggests an algorithm that performs efficiently with small values of $K$, $m$, and $n$ but may become substantially more resource-intensive as these parameters increase. The number tours $T$ indicates that the entire complex operation is executed $T$ times, further amplifying the effect of the inner complexities.

Optimizing such an algorithm involves minimizing the repetition count ($T$), reducing the complexity of operations dependent on $K$, $m$, and $n$.

\subsubsection{Time Complexity Analysis of the Genetic Algorithm for TSP}
\label{ga_tsp}

When $K > 6$, we obtain 100 tours from a heuristic TSP solution with the Genetic Algorithm (GA), which addresses optimization problems by evolving a population of candidate solutions through selection, crossover (recombination), and mutation over multiple generations, aiming to improve the fitness of solutions.

The time complexity of applying Genetic Algorithms to solve the TSP is influenced by several key factors, including the number of nodes (\(K\)), population size (\(P\)), number of generations (\(G\)), and the complexities of the crossover (\(C(K)\)), mutation (\(M(K)\)), and selection (\(S(K)\)) operations.

Given these parameters, we can express the overall time complexity as:
\begin{itemize}
	\item $O(G \cdot (P \cdot (S(K) + P \cdot (C(K) + M(K)))))$
\end{itemize}

{\bf Detailed Explanation}
\begin{itemize}
	\item The selection operation's complexity, \(S(K)\), may depend on evaluating the fitness of each solution, which in turn relies on computing the total distance of a path through \(K\) nodes.
	\item The complexities of the crossover and mutation operations, \(C(K)\) and \(M(K)\), are functions of \(K\) because they involve manipulating sequences representing paths through the \(K\) nodes.
	\item Each generation involves selecting solutions based on fitness, generating new solutions through crossover, and applying mutations, leading to the multiplication by \(G\).
\end{itemize}

{\bf Impact of the number of nodes (\(K\))}
\begin{itemize}
	\item The direct involvement of \(K\) in the complexities of \(C(K)\), \(M(K)\), and \(S(K)\) highlights how the number of nodes affects the algorithm's performance. Specifically, operations that manipulate or evaluate paths become more complex as \(K\) increases.
	\item The encoding of solutions and the calculation of fitness values, essential for the selection process, inherently depend on \(K\), emphasizing its role in the GA's computational demands.
\end{itemize}

}

@@

\begin{algorithm}[H]
	\caption{Generating items}  \label{alg:itemsgen}
	\begin{algorithmic}[1]
		\State {\bf ItemsGeneration} \,\, {\it in: $scenario, surplus$} \, \, {\it out: $N$}
		\State Let $L$ be the set of nodes and $M$ the set of pallets  \label{ig:LM}
		\State $limit \gets surplus \times \sum_{i=1}^{m} V_i$ \label{ig:extended}
		\For{$k \gets 0$ to $K$}
		\State $N_k \gets \varnothing$
		\State $j \gets 0$
		\State $vol \gets 0$	\label{ig:totals}	
		\While{$vol < limit$}
		\State $j \gets j+1$
		\State Let $t_j^k$\/ be the item $j$\/ at the node $k$
		\Repeat
		\State $to_j \gets RandomInt(0, K)$ \label{ig:dest}
		\Until{$to_j \neq k$}
		\State $x = Roulette(item)$ biased through $P$ \label{ig:weight1} \Comment{From Table \ref{tab:weights}}	 
		\State $w_j \gets RandomInt(low(x), high(x))$        \label{ig:weight2}		
		\State $s_j \gets \round{100 \times (1 - \log_{10}(RandomInt(1, 9)))} $ \label{ig:score}
		\State $v_j \gets w_j / RandomInt(148, 344)$ \label{ig:volume}
		\State $vol \gets vol + v_j$ 
		\State $N_k \gets N_k \cup \{t_j^k\}$ 
		\EndWhile
		\EndFor
		\State $N \gets \bigcup_{0 \leq k \leq K} N_k$
	\end{algorithmic}
\end{algorithm}

@@

We will first present the results obtained in practical cases, that is, when $1<K\leq6$. Next, we will show how the same {\it Shims}\/ heuristic remains robust when $K>6$.

@@

These last results that we will show here do not correspond to practical cases of air transport, as tours where $K>6$ very rarely occur. However, it is possible to see that the {\it Shims} heuristic maintains robust behaviour as the number of nodes grows, that is, it could be adapted to similar contexts
({\color{red}vehicles and warehouses} {\color{blue}ships and trucks})
, for example), where there may be more nodes.

>>>>

\begin{abstract}
In the aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of load imbalance due to the urgency required for loading, immediate take-off, and mission accomplishment. Transport planners deal with trip itineraries, prioritisation of items, building up pallets, and balanced loading, but there are no commercially available systems that can integrally assist in all these requirements. This enables other risks, such as improper delivery, excessive fuel burn, and possible safety issues due to cargo imbalance, as well as a longer than necessary turn-around time. This {\it NP-hard}\/ problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modelled using standardised pallets in fixed positions. We developed a strategy to solve this problem, considering historical transport data from some Brazilian hub networks, and performed several experiments with a commercial solver, five known meta-heuristics, and a new heuristic designed specifically for this problem. By using a portable computer, our strategy quickly found practical solutions to a wide range of real problems in much less than operationally acceptable time.
\end{abstract}

@@

@@

The problem addressed in this work is the definition of the route of an aircraft that loads and unloads hundreds of items in several hubs, dealing with weight, volume, and balance constraints, and maximising the benefit-cost ratio. This challenge has become even more acute during the COVID-19 pandemic due to the need for urgent medical supplies. Delays in shipping essential items such as respirators to critical areas highlighted the need for optimised solutions.

@@

		\caption{Distances between some Brazilian airports ($km$)} \label{tab:costs}
		\centering
		\footnotesize

@@

In all these works, there is a great diversity of scenarios and solutions: some consider items in two dimensions, and others in three dimensions; some used integer programming, and others developed specific heuristics. Among the most recent, we can highlight:

@@

	\item \cite{zhao2023} presented three models that use integer programming for air cargo planning and weight balance optimization: bi-objective optimization (BOM), combinatorial optimization (COM), and enhanced combinatorial optimization (IOM). Considering a Boeing 777F in several scenarios, the tests revealed performance problems: BOM is fast, but produces large CG deviation; COM offers accurate optimization, but with impractical runtimes; IOM provides a balanced solution, improving speed over COM, but requiring high computational demands in some cases. Although IOM stands out for its effectiveness, all models face trade-offs between speed, accuracy and computational efficiency. This work alerted us to potential performance issues in solution methods.

@@

Throughout this text, we call {\it packed content}\/ (see Figure \ref{fig:packed}) a set of items of the same destination stacked on a pallet and covered with a restraining net. It is considered a single item, having the same attributes as its components, whose values are the sum of individual scores, weights, and volumes. To ensure accuracy in pickup and delivery operations, packed content must remain on board until its destination.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.3]{packed.jpg}
	\caption{A {\it packed content}\/ on 463L pallet inside a {\it Boeing C-17}}
	\small\textsuperscript{Source: From Wikimedia Commons, the free media repository}
	\label{fig:packed}
\end{figure}

@@

\section{Solution process}
\label{algorithms}

Throughout our research, we have thoughtfully described ACLP+RPDP in standard MIP format and found that no solver can handle its practical cases in a feasible time. Thus, as ACLP+RPDP is highly complex, involving four intractable and interconnected sub-problems, we decided to focus on {\it real cases}, developing quick node-by-node solutions, not necessarily optimal, but which would allow us to obtain a complete tour.

@@

			\node [mybox, below=0.7cm of one] (two) { \small {\bf Number of tested tours:}\\
				\small \ \ \ {{\bf if} $K \leq 6$}\\
				\small \ \ \ \ \ \ {$ntours = 2$ \ \ (optimal TSP solutions)}\\
				\small \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\it or}\\				
				\small \ \ \ \ \ \ {$ntours = K!$ \ \ (all possible tours)}\\
				\small \ \ \ {\bf else}\\
				\small \ \ \ \ \ \ {$ntours = 100$ \ \ (tours obtained with a TSP heuristic)}
			};

@@

\begin{algorithm}[htp]
	\small
	\caption{Solving ACLP+RPDP}  \label{alg:main}

@@

				\For {$i \gets 1$ to $ntours$} \label{main:loop1}

					\State $f_i \gets SolveTour(\pi[i], L, M, C, N, method, tmax/ntours)$ \label{main:fpi}

@@

				\State $answer[scenario,surplus,method] \gets \max f_i$ \label{main:a2}
			\EndFor
	\end{algorithmic}

@@

$\pi[\ ]$ is a vector of tours indexed from 1 to $ntours$. When $ntours=2$, $\pi[1]$\/ and $\pi[2]$\/ are the optimal solutions of the corresponding TSP. When $ntours=K!$, $\pi[i]$\/ is the $i^{th}$ permutation of $S_K$. Finally, when $ntours=100$, $\pi[1], \ldots, \pi[100]$\/ are the solutions provided by a TSP heuristic.

@@

\begin{algorithm}[H]
	\small
	\caption{Solving tour $\pi$ with $method$}  \label{alg:tour}
	\begin{algorithmic}[1]

@@

When the aircraft is at the base, the initial graph $G_1$\/ is empty, and there are no packed contents (line \ref{tour:g11}). Otherwise, $UpdatePacked$\/ (line \ref{tour:dest}) returns the set of packed contents that have not yet reached their destination and remain on board, rearranging them on the pallets to minimize CG deviation. This allocation is stored in graph $G_1$\/ (line \ref{tour:g12}).

@@

\begin{algorithm}[H]
	\small
	\caption{Updating the packed contents that remain boarded at node $\pi_k$}  \label{alg:cons}
	\begin{algorithmic}[1]

@@

\begin{algorithm}[H]
	\small
	\caption{Setting pallets destination based on the items to be embarked at node $\pi_k$}  \label{alg:dest}
	\begin{algorithmic}[1]

		\State {\bf SetPalletsDestinations} \,\, {\it in: $M, \pi_k$} \, \, {\it out: $M$}
		\State $vol_x \gets  0$, $x \in L_{\pi_k}$  \label{dest:vector1}
		\State $max \gets 0$  \Comment{destination with maximum volume demand}
		\State $total \gets 0$ 
		\For{$j \gets 1$ to $n_{\pi_k}$}
		\If {$to_j \in L_{\pi_k}$} 
		\State $vol_{to_j} \gets vol_{to_j} + v_j$ 
		\State $total \gets total + v_j$ 
		\If {$vol_{to_j} > vol_{max}$}
		\State $max \gets to_j$ \label{dest:max} 
		\EndIf
		\EndIf
		\EndFor
		\For{$x \in L_{\pi_k}$} \label{dest:propor1}
		\If {$vol_{x} \neq 0$}
		\State $needed \gets \max \{1, \lfloor{ (m-m_{\pi_k}) \times vol_{x}/total}\rfloor \}$  
		\label{dest:needed}
		\State $np \gets 0$
		\For{$i \gets 1$ to $m$}
		\If{($np < needed$) {\bf and} ($T_i^{\pi_k} = -1$)}  
		\State $T_i^{\pi_k} \gets x$
		\State $np \gets np + 1$ \Comment{number of necessary pallets to node $x$}
		\EndIf
		\EndFor
		\EndIf 
		\EndFor \label{dest:propor2}
		\For{$i \gets 1$ to $m$} \label{dest:max1}
		\If{$T_i^{\pi_k} \gets -1$} 
		\State $T_i^{\pi_k} \gets max$  \Comment{any remaining pallet is assigned to the maximum demand destination}
		\EndIf
		\EndFor \label{dest:max2}

@@

\begin{algorithm}[H]
	\small
	\caption{Cargo score and aircraft torque}  \label{alg:eval}
	\begin{algorithmic}[1]

@@

In this way, the mathematical model for $SolveNode(MIP,\pi_k,G, tmax)$\/ becomes simpler, which finds an allocation of available items at node $\pi_k$\/ using previously defined values of $L_{\pi_k}$, $T_i^{\pi_k}$, and $a^{\pi_k}_q$. Thus, we use a MIP solver with a runtime limit $tmax$\/ at node $\pi_k$ to maximize the objective function (\ref{eq:obj2}) with the calculus equations (\ref{eq:newf}) to (\ref{eq:costs2}), subject to the constraints (\ref{eq:tau2}) to (\ref{eq:ifB}). The binary variables $X_{ij}$\/ and $Y_{iq}$\/ define the sets of edges $E_{N_{\pi_k}}$\/ and $E_{Q_{\pi_k}}$, respectively, included in graph $G$.

@@

The selection of edges for $E_{ N_{\pi_k} }$ uses the {\it edge attractiveness}\/ $\theta_{ij}$\/ (\ref{eq:edge}), which can be understood as the tendency to allocate item $j$\/ to pallet $i$\/ at node $\pi_k$. It is directly proportional to the score, and inversely to the volume and the torque of each item.

@@

Considering only the items that can be shipped at node $\pi_k$, Figure \ref{fig:whip} represents $n_{\pi_k}$\/ possible edges $e_{ij}$\/ of pallet $i$\/ sorted by $\theta_{ij}$\/ in non-ascending order. Initially, {\it Shims}\/ builds a greedy solution for pallet $i$\/ selecting edges up to index $\eta_1$ ({\it greedy phase}). Then, with the edges between $\eta_1$\/ and $\eta_2$, it elaborates different possible complements ({\it composition phase}), including later the best ones in the same pallet ({\it selection phase}). {\it Shims}\/ is depicted in Algorithm \ref{alg:shims}.

@@

		\State Let $E$\/ be an array of $n_{\pi_k}$ possibles edges of pallet $i$\/ sorted by $\theta_{ij}$\/ in non-ascending order \label{shims:possible}
		\State $\eta_1 \gets 1$  
		\Repeat
		\State $e_{ij} \gets E_{\eta_1}$

@@

$tmax$\/ is the runtime limit for {\it Shims}, and $level_1$\/ and $level_2$\/ are volume thresholds for indices $\eta_1$ and $\eta_2$, respectively. Initially, $Q_{\pi_k}$\/ (line \ref{shims:initQ}) corresponds to the packed contents that remain on board. It is important to remember that $E_{Q_{\pi_k}}$\/ and $M$\/ were modified by the procedure $UpdatePacked(M, Q_{\pi_k}, \pi_k)$ and the procedure $SetPalletsDestinations(M, \pi_k)$. Then, the pallets are considered in non-descending order of $|D_i^{long}|$.

@@

	\item In the {\it greedy phase}\/ (lines \ref{shims:pallets}-\ref{shims:endfirst}), a partial solution for each pallet $i$\/ is constructed by adding edges following $\theta_{ij}$\/ order. Indices $\eta_1$\/ and $\eta_2$\/ refer to the accumulated volumes $V_i \times level_1$\/ and $V_i \times level_2$, respectively, which were defined empirically by the {\it irace}\/ tool \citep{LopezIbanezManuel2016}. We will explain this in subsection \ref{items}.

@@

\subsection{Time complexity using {\it Shims}}

We finish this section with the analysis of the time complexity of our ACLP+RPDP solution, considering the use of the {\it Shims}\/ heuristic.

In this process, when $K>6$\/ (that is, in unusual cases of air transportation), we need a GA-based TSP heuristic to generate 100 tours of size $K$. This heuristic can be chosen and calibrated to be fast enough: specifically, in the tests we will present in the next section, it took just 33 seconds with $K=15$. For this reason, we will not perform its complexity analysis. We will also not consider the time spent by the {\it irace}\/ tool, in defining parameters for {\it Shims}, as it is only executed once in the calibration of our method.

Let $n = \max n_k$. Without loss of generality, we assume that $m = \mathcal{O}(n)$ and $\max m_k = \mathcal{O}(n)$. Therefore, to read the input data, Algorithm \ref{alg:main} requires time $\mathcal{O}(m + K^2 + K*n + m*\max m_k) = \mathcal{O}( n*m)$, considering $K^2 = \mathcal{O}(m*n)$. Since $ntours \leq 6!$, it is enough to calculate the complexity of Algorithm \ref{alg:tour}.

	\item Its initial variables are $\mathcal{O}(K*n)$.

	\item Algorithm \ref{alg:cons} minimizes CG deviation and updates pallets.
	We can consider that the MIP solver spends time $\mathcal{O}(m^2) = O(m*n)$, as it needs to allocate at most $m$\/ packed contents to $m$\/ pallets. The final loop of Algorithm \ref{alg:cons} has the same complexity.

	\item Algorithm \ref{alg:dest} takes $\mathcal{O}(n+K+m)$ time.

	\item {\it Shims}, described in Algorithm \ref{alg:shims}, spends $\mathcal{O}(m*\log m)$ in ordering the pallets and then performs a loop with $m$\/ iterations. Each of these iterations spends $\mathcal{O}(n*\log n)$\/ on ordering the edges by the value of $\theta_{ij}$, and $\mathcal{O}(n)$\/ on the others statements. So its total time is $\mathcal{O}(m*n*\log n)$.

	\item Algorithm \ref{alg:eval} takes $\mathcal{O}(m*n)$ time.

Therefore, we can conclude that the total time of Algorithm \ref{alg:tour} is $\mathcal{O}(K*m*n*\log n)$, which is the time complexity of the ACLP+RPDP solution process .

@@

\begin{algorithm}[H]
	\small
	\caption{Generating items}  \label{alg:itemsgen}
	\begin{algorithmic}[1]
		\State {\bf ItemsGeneration} \,\, {\it in: $scenario, surplus$} \, \, {\it out: $N$}
		\State Let $L$ be the set of nodes and $M$ the set of pallets  \label{ig:LM}
		\State $limit \gets surplus \times \sum_{i=1}^{m} V_i$ \label{ig:extended}
		\For{$k \gets 0$ to $K$}
		\State $N_k \gets \varnothing$
		\State $j \gets 0$
		\State $vol \gets 0$	\label{ig:totals}	
		\While{$vol < limit$}
		\State $j \gets j+1$
		\State Let $t_j^k$\/ be the item $j$\/ at the node $k$
		\Repeat
		\State $to_j \gets RandomInt(0, K)$ \label{ig:dest}
		\Until{$to_j \neq k$}
		\State $x = Roulette(item)$ biased through $P$ \label{ig:weight1} \Comment{From Table \ref{tab:weights}}	 
		\State $w_j \gets RandomInt(low(x), high(x))$        \label{ig:weight2}		
		\State $s_j \gets \round{100 \times (1 - \log_{10}(RandomInt(1, 9)))} $ \label{ig:score}
		\State $v_j \gets w_j / RandomInt(148, 344)$ \label{ig:volume}
		\State $vol \gets vol + v_j$ 
		\State $N_k \gets N_k \cup \{t_j^k\}$ 
		\EndWhile
		\EndFor
		\State $N \gets \bigcup_{0 \leq k \leq K} N_k$
	\end{algorithmic}
\end{algorithm}

@@

We will first present the results obtained in practical cases, that is, when $1<K\leq6$. Next, we will show how {\it Shims}\/ remains robust when $K>6$.

@@

These last results do not correspond to practical cases of air transport, as tours where $K>6$ very rarely occur. However, it is possible to see that the {\it Shims}\/ maintains robust behaviour as the number of nodes grows, that is, it could be adapted to similar contexts (ships and trucks, for example), where there may be more nodes.
